\plaintitle

### 2016-05-21 paper/paper.tex
Reworked related work

\todo{Rework realted work}
\section{Related Work}\label{sec:literature}

On the IEEE Visualization Conference in 2013, the first workshop on sports data
visualization was held \cite{ieeevis}. There is shown that the area of sport
produces large amounts of data, which can be used for building visualizations
and making analyses of the data. In the past visualizations of sports existed
already, but there were not a lot of visualizations, and they were simple in
the sense that they just represented the data without trying to give insights
or to approach the data on a different way. There are some, but limited,
sports-specific visualizations, like the basketball shot chart. This is a
visualization frequentely seen when looking for visualizations about the NBA.
Some shot chart vizualisations try to be different than others, but the general
design and idea stays roughly the same. 
% 2 goeie papers die hierin verwerkt kunnen worden
%http://www.sloansportsconference.com/wp-content/uploads/2012/02/Goldsberry_Sloan_Submission.pdf
%http://vis.berkeley.edu/courses/cs294-10-sp10/wiki/images/archive/a/a2/20100511091342!Stephen_Chu_-_Information_Visualization_in_the_NBA_-_The_Shot_Chart_16.pdf

With our visualization we try to make it easier to interpret a collection of
data about the NBA, and try to let users find patterns and insights in the
data. A same kind of incentive was the start for \cite{cox2006sportsvis} who
created a visualization to help people discover meaning in statistics generated
during sporting events in the case of baseball.

(From bubble view)
A more traditional visualization of a play-off is the team double elimination
bracket.  Although this is a more common representation , and there are some
nice desings like in this NBA 2012 play-offs
visualisation~\cite{tournamentladder} where a sunburst visualisation is used,
we wanted to create a more compact version by eliminating the recurring
representation of a team. We considered enclosed circles, as used in this
visualisation of the wimbledon tournament roster~\cite{enclosedcircles}.
Although the enclosed circles is a compact representation as well, we have
continued with our own design as we believe this view is more original and
innovative. Furthermore we are convinced that our view gives a good
illustration of the end positions. It's generally assumed that the team that is
on top of the ``ladder'' is the winning team. We didn't encouter a similar view
when searching for alternatives to visualize a play-off or championship. A new
visualization has the advantage that it will trigger users as well.

(From field view)
One good alternative was the shooting signature.\cite{peterbeshai} However,
creating such a visualization is again a project on itself and due to time
constraints we left it aside. Other possibilities such as a heatmap or spider
web didn't fulfill our need to give the visitor a clear overview of how a player
performed compared to its peergroup.

@@

\subsection{Technology}
To create the visualization, the d3 javascript framework~\cite{d3.js} was used
in combination with html5~\cite{html5} and jquery~\cite{jquery}.  Note that D3
is based on HTML, CSS, Javascript and SVG. This allows easy access to the visualization as
most modern browsers are capable of handling these webstandards.  The choice not
to use the d3 framework for the entire visualization was made to ease the layout
configuration of the visualization. Instead, html was used to do the global
layout of the visualization. We opted to structure the site with multiple divs
as containers. To each container we allocate a svg with a specific
visualization.

>>>>

\section{Related Work}\label{sec:literature}
Recently, there has been an increase in the interest in visualizations of sports
data. This is best exemplified by theworkshop on sports data visualization at
the IEEE Visualization Conference in 2013\cite{ieeevis}.

There are a number of traditional basketball visualizations. One of the most
popular ones is the shot chart. This visualizes the shots of a player or team
around the hoop. It provides insight into the shooting ability of a
player\cite{goldsberry,stephenchu}.

This view features heavily in Peter Beshai's Buckets\cite{peterbeshai}. This is
an extensive visualization of NBA player data. It provides not only the player's
shot chart for various years, but his shooting signature and several line graphs
as well. It is strongly focused on individual player's shooting abilities.

Visualizations focusing on basketball teams and their performance are
significantly more difficult to find. For the playoffs, there is the traditional
playoffs bracket~\cite{tournamentladder}. This bracket structure has the
advantage of being well known and as such, not requiring much explanation. It is
flawed however. It duplicates much information by displaying teams multiple
times.

An interesting alternative is the sunburst tournament
visualization\cite{sunburst}. As the name suggests, this visualizes the
tournament as a sunburst, with the winning team taking the part of the circle
closer to the center. This does not eliminate the problem of the recurring teams
however.

For the wimbledon tennis tournament, an enclosed circles visualization was
created~\cite{enclosedcircles}. This visualization eliminates the repeating
occurrences of the same team. However, it wastes a lot of space.

England's Permier League results were visualized by Ami Sedghi for the
Guardian\ref{premierleague}. This visualization shows the team's rankings as
line graphs and allows highlighting of individual teams.

Tan et al. created an adaption of the traditional tournament bracket to better
support moden interaction\cite{adaptivitree}. The tool, AdaptiviTree, changes
the brackets shape and adds colored lines to show available information as the
user pciks his own bracket.

An alternative to the traditional bracket structure can be found in the
visualization of the race to the white house in 2012 made for the New York Times
by Mike Bostock and Shan Carter\cite{whitehousepath}. They use a tree-like
structure where each split represents a state being one by one candidate or the
other. The interactivity is very important here as it allows users to highlight
one possible path out of many.

The entire history of the NBA and the best teams were analyzed by
fivethirtyeight~\cite{fivethirtyeight}. They calculated an elo score for all
teams an visualized them on line graphs.

Pagno et al. provide a number of different visualizations\cite{starplots}. The
most interesting one is the starplot. This plot several statistics for a team on
an axis pointing radially outward from a common center. These points are then
connected. It allows the user to get basic information about a team or player at
a glance. Because of this speed of conveying information, it is also a prevalent
visualization in sports games.

@@

\subsection{Technology}
To create the visualization, the d3 javascript framework~\cite{d3} was used in
combination with html5~\cite{html5} and jquery~\cite{jquery}.  Note that D3 is
based on HTML, CSS, Javascript and SVG. This allows easy access to the
visualization as most modern browsers are capable of handling these
webstandards.  The choice not to use the d3 framework for the entire
visualization was made to ease the layout configuration of the visualization.
Instead, html was used to do the global layout of the visualization. We opted to
structure the site with multiple divs as containers. To each container we
allocate a svg with a specific visualization.
